Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-16 paper/structureScene.tex
Bioacoustics -> ecoacoustics

An important task in computational bioacoustics and related fields, such as ecouacoustics, is to determine the similarity of two audio scenes. This is often done using bag-of-frames models and other representations based on summary statistics. While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background. To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure. To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs). Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.

@@

% Note that keywords are not normally used for peerreview paper.
\begin{IEEEkeywords}
bioacoustics, ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
\end{IEEEkeywords}

>>>>

un important task in computational ecoacoustics is to determine the similarity of two audio scenes. This is often done using bag-of-frames models and other representations based on summary statistics. While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background. To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure. To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs). Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.

@@

% Note that keywords are not normally used for peerreview paper.
\begin{IEEEkeywords}
ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
\end{IEEEkeywords}
